Efficiently summarising relationships in large samples: a general duality between statistics of genealogies and genomes

### 2019-09-06 treestats_paper.tex
Say why trees only partially known
Add comment on older refs.

The trees describing genealogical relationships as one moves along the genome
-- also described by the Ancestral Recombination Graph \citep[][or ARG,]{arg} --
contain a good deal of information about the history and ongoing dynamics of a population.
Given sufficiently many mutations, it would be possible to infer the genealogical tree
at each point along the genome \citep{felsenstein_book},
and each mutation gives a small piece of information about the tree at that site
\citep{semple2003phylogenetics}.
However, population genetics has the task of summarizing relationships across \emph{many} trees,
each only partially known.
The basic relationship between tree shape and summaries of genetic variation
is that if mutations are neutral then the expected number that occur on a given branch
is proportional to the length of that branch,
thus establishing a duality between genealogical and genetic summaries,
under the infinite-sites model of mutation \citep{ralph2019empirical}.

>>>>

The trees describing genealogical relationships as one moves along the genome
-- also described by the Ancestral Recombination Graph \citep[][or ARG,]{arg} --
contain a good deal of information about the history and ongoing dynamics of a population.
Given sufficiently many mutations, it would be possible to infer the genealogical tree
at each point along the genome \citep{felsenstein_book},
and each mutation gives a small piece of information about the tree at that site
\citep{semple2003phylogenetics}.
However, population genetics has the task of summarizing relationships across \emph{many} trees,
each only partially known because mutations are sparse in the data.
The basic relationship between tree shape and summaries of genetic variation
is that if mutations are neutral then the expected number that occur on a given branch
is proportional to the length of that branch,
thus establishing a duality between genealogical and genetic summaries,
under the infinite-sites model of mutation \citep{ralph2019empirical}.
\krt{The idea is much older than this: Taiima 83, 89(?), Fu 95 are all relevant here.  The latter
derives a lot of properties of the expectations of "node weights" under Kingman's coalescent, including
the polarised vs non- cases.}
